---
title: "Final Project"
author: "Haichen Wei"
date: "Dec 13, 2020"
output: html_notebook
---

# Which is a Better Choice for Breakfast, Cereals, Starbucks or McDonald's?

## Introduction

This project will compare the nutrition facts cereal products, Starbucks and McDonald's menu, and tell people which one is better to choose as breakfast. This analysis uses three dataset, all of them come from Kaggle. These datasets provide some basic nutrition facts of food, such as Calories, Fat, Protein and so on. Each dataset has different numbers of columns, so this analysis only choose the common columns of them.

## Importance of Exploring Nutrition Facts

Lots of people need to go to work or school in the morning. If they don't get up early, they don't have much time to prepare meal. At this moment, fast food becomes a easy way to deal with the breakfast. Cereal is the most common products for breakfast, you can easily make it at home. The Starbucks and McDonald's are also popular choices, people can buy something ready to eat there. Fast food can save time, however, we also need to eat healthy. This project helps to analyze the nutritions of three fast food, and summary which fast foof is better.

## Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: 
library(tidyverse)
library(mosaic)
library(DataComputing)
library(dplyr)
library(pryr)
```

## Data Access

The description of 80 Cereals says it contains 80 kinds of cereals, but the dataset only have 77 cases in total, and it has 16 columns. Starbucks contains 113 cases and 6 columns. McDonalds has 260 cases and 24 columns. Since each dataset has different number of columns, so I choose the variables in common. Three datasets have 6 common variables, name, Calories, Fat, Carb., Fiber, Protein. 

### Cereals
```{r}
# read the dataset
Cereals <- read.csv("/Users/Zoey/Desktop/cereal.csv")
# inspect data
glimpse(Cereals)
```
### Starbucks
```{r}
# read the dataset
Starbucks <- read.csv("/Users/Zoey/Desktop/food.csv")
# inspect data
names(Starbucks)
nrow(Starbucks)
```
### McDonald's
```{r}
# read the dataset
McDonalds <- read.csv("/Users/Zoey/Desktop/menu.csv")
# inspect data
str(McDonalds)
```
## Data Wrangling

### Cereals
```{r}
# select common variables and rename it, add a type colomn
Cereals <-
  Cereals %>%
  mutate(Type = "Cereal") %>%
  select(name, Type, calories, protein, fat, fiber, carbo) %>%
  rename(Item = `name`, 
         Calories = `calories`, 
         Protein = `protein`, 
         Fat = `fat`, 
         Fiber = `fiber`, 
         Carb. = `carbo`)
Cereals
```

### Starbucks
```{r}
# add a type column to Starbucks and rename the variables
Starbucks <- 
  Starbucks %>%
  mutate(Type = "Starbucks") %>%
  rename(Item = `item`, 
         Protein = `Protein..g.`, 
         Fat = `Fat..g.`, 
         Fiber = `Fiber..g.`, 
         Carb. = `Carb...g.`)
Starbucks
```

### McDonald's

```{r}
#spread category 
McDonalds %>%
  group_by(Item, Category) %>%
  summarise(total = sum(Calories))%>%
  spread(key = Category, value = total, fill = 0)
```

```{r}
#find the item which has egg or sausage
McDonalds %>%
  filter(grepl("Sausage|Egg",Item))
```

```{r}
# select common variables and add a type colomn
McDonalds <-
  McDonalds %>%
  filter(grepl("Breakfast", Category)) %>% #McDonald's has a categore column, so I only choose the breakfast food, and the total cases for breakfast is 42.
  mutate(Type = "McDonalds") %>%
  select(Item, Type, Calories, Total.Fat, Carbohydrates, Dietary.Fiber, Protein) %>%
  rename(Fat = `Total.Fat`, 
         Fiber = `Dietary.Fiber`, 
         Carb. = `Carbohydrates`)
McDonalds
```

```{r}
# combine the table
Food <- rbind(Cereals, Starbucks, McDonalds)
Food
```

After calculating the range and mean for each variable, McDonald's has the most calories, protein, fat, fiber and carbohydrates. Starbucks is the second, and cereal is the third.
```{r}
Range <- 
  Food %>%
  group_by(Type) %>%
  summarise(range_cal=max(Calories)-min(Calories), 
            range_protein=max(Protein)-min(Protein),
            range_fat=max(Fat)-min(Fat),
            range_fiber=max(Fiber)-min(Fiber),
            range_carb=max(Carb.)-min(Carb.))
Range
Mean <-
  Food %>%
  group_by(Type) %>%
  summarise(mean_cal=mean(Calories), 
            mean_protein=mean(Protein),
            mean_fat=mean(Fat),
            mean_fiber=mean(Fiber),
            mean_carb=mean(Carb.))
Mean
```

```{r}
# gather the table
Mean <-
Mean %>%
  gather(key = variables, value = mean, mean_cal, mean_protein, mean_fat, mean_fiber, mean_carb)

Range <-
Range %>%
  gather(key = variables, value = range, range_cal, range_protein, range_fat, range_fiber, range_carb)
```

```{r}
#join the tables
Join <-
Range %>%
  full_join(Mean) 
```

## Data Visualization

The scatter plot graph shows McDonald's has the highest calories and fat overall. The most plots of Cereal are between 0 to 150, it has very low calories and fat. The plots of Starbucks are between 0 to 750, so the food is under 750 calories and 40 fat.The plots of McDonald's covered from 150 to 1200 calories and 0 to 60 fat.

```{r}
ggplot(data=Food,aes(x=Calories,y=Fat ))+geom_point()+ facet_wrap(~Type,ncol=4) + stat_smooth(method = 'loess')
```

The following bar graph shows calorie has the largest range in the food, and the values of fiber are very small, it is almost covered by other variables.

```{r}
ggplot(data=Range,aes(x=Type,y=range ,fill=variables))+geom_bar(stat='identity',position='stack', width=.9) 
```

The following overlaid density distributions show that the food in Starbucks and McDonald's has larger carbohydrates than cereals, and the fiber in these three types of food is very little.

```{r}
ggplot(data=Food,aes(x = Carb., color=Type)) + geom_density() + geom_density(data=Food,aes(x = Fiber, fill=Type))
```
## Technical Challenge

The biggest technical challenge I encountered during the investigation is how to load the dataset from Kaggle. At first, I tried to use read.csv to read the file, but it came out a table of strings. I also tried read.file, and it didn't work. Then, I found that we can down the csv file directly from the Kaggle website,so I download the csv file to my laptop and use read.csv to read the file.

## Conclusion

All of these three types of food are not very healthy, they has little fiber and high calories, fat, carbohydrates and protein. If people eat these fast food every day, they will get fat quickly. Cereals is the safest choice if you don't want to get fat, but if you don't worry about the weight, the food in Starbucks and McDonald's can make you full.

## Reference

Crawford, C.(2017). 80 Cereals. Retrieved from https://www.kaggle.com/crawford/80-cereals?select=cereal.csv.

Starbucks.(2017). Nutrition facts for Starbucks Menu. https://www.kaggle.com/starbucks/starbucks-menu?select=starbucks-menu-nutrition-food.csv

McDonald.(2017). Nutrition Facts for McDonald's Menu. https://www.kaggle.com/mcdonalds/nutrition-facts?select=menu.csv

## GitHub Links

https://hxw5245.github.io/STAT184FinalProject/

https://github.com/hxw5245/STAT184FinalProject.git

